---
title: "Using APIs: Creating Vignette on Stock Data as Example"
author: "Evan Brown"
date: '2022-06-23'
output: 
  github_document:
    toc: true
    html_preview: false
---

```{r global options, include=FALSE}
knitr::opts_chunk$set(fig.path = "README_files/figure-gfm")
```

This is a vignette to show how to retrieve data from an [API](https://aws.amazon.com/what-is/api/#:~:text=API%20stands%20for%20Application%20Programming,other%20using%20requests%20and%20responses.). To demonstrate, I'll be working with the [Polygon Financial API](https://polygon.io/docs/stocks/getting-started). For the purposes of this work, I used a free account with Polygon, limiting my data retrieval "from-date" to 2 years. 

In this vignette, I will work to pull data from the [S&P Oil & Gas Exploration & Production Industry Index](https://www.spglobal.com/spdji/en/indices/equity/sp-oil-gas-exploration-production-select-industry-index/#data) and compare endpoints of some top index constituents to industry [ETF's](https://www.investopedia.com/terms/e/etf.asp#:~:text=An%20exchange%2Dtraded%20fund%20(ETF)%20is%20a%20type%20of,that%20a%20regular%20stock%20can.) and see if I can find any similarities or differences over the past 2 years. 


# **Required Functions**

These R packages are required in order to work with the Polygon API:

- [tidyverse](https://www.tidyverse.org/packages/)  
- [jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) 
- [httr](https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html)  

```{r, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
#Reading in packages
library(tidyverse)
library(httr)
library(jsonlite)
```

# **Functions for API Interaction**

This portion of the vignette will show where I define functions for interacting with the Polygon Financial API.

## 'marketcap' Function

This function returns the [market cap](https://www.investopedia.com/terms/m/marketcapitalization.asp) for the user provided ticker and date. apikey default is my key, but users of the function should pass their own key in the function call options. User keys can be generated [here](https://polygon.io/).

```{r, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

marketcap <- function(ticker, date, apikey = "OrlbxnjeCyqGDGkKtpIqxKKs0f8Eh77C"){
  
  #Set URL partitions
  baseURL <- "https://api.polygon.io/v3/reference/tickers/"
    ticker <- ticker
      dateadjustment <- "?date="
        date <- date
          adjustments <- "&apiKey="
            apikey <- apikey
          
  #Paste partitions to get full URL with desired ticker
  fullURL <- paste0(baseURL, ticker, dateadjustment, date, adjustments, apikey)   
  
  #Get api data using get function
  api_data <- GET(fullURL)
    #Convert to character string with fromJSON function
    parsed_api_data <- fromJSON(rawToChar(api_data$content))
      #Save as R object
      parsed_api_data_results <- parsed_api_data$results
      results <- as_data_frame(parsed_api_data_results[names(parsed_api_data_results) != "address"])
      ticker <- results$ticker
      marketcap <- results$market_cap
      year <- substr(date, 1, 4)
    
      finaldata <- as_data_frame(cbind(ticker, marketcap, year))
      finaldata <- finaldata[1,]
      
      return(finaldata)
}

```

## 'aggregate' function

This function returns daily stock endpoints from 06-21-2021 to 06-22-2022 as a tibble for the user provided tibble. apikey default is my key, but users of the function should pass their own key in the function call options. User keys can be generated [here](https://polygon.io/).

```{r, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

aggregate<- function(ticker, apikey = "OrlbxnjeCyqGDGkKtpIqxKKs0f8Eh77C"){
  
  #Set URL partitions
  baseURL <- "https://api.polygon.io/v2/aggs/ticker/"
    ticker <- ticker
      adjustments <- "/range/1/day/2020-06-25/2022-06-25?adjusted=true&sort=asc&limit=5000&apiKey="
        apikey <- apikey
  
  #Paste partitions to get full URL with desired ticker
  fullURL <- paste0(baseURL, ticker, adjustments, apikey)
  
  #Get api data using get function
  api_data <- GET(fullURL)
    #Convert to character string with fromJSON function
    parsed_api_data <- fromJSON(rawToChar(api_data$content))
      #Save as R object
      parsed_api_data_results <- parsed_api_data$results
        #Convert to tibble
        parsed_tibble <- as_tibble(parsed_api_data_results)

  #Create sequence of raw dates
  dates <- (seq(as.Date("2020-06-25"),as.Date("2022-06-25"),by = 1))
    #Filtering weekends (market not open)
    dates <- as.tibble((dates[!weekdays(dates) %in% c("Saturday","Sunday")]))
      #Filtering holidays (market not open)
        holidaydates <- as.tibble(as.Date(c("2020-01-01", "2020-01-20", "2020-02-17", "2020-04-10", "2020-05-25", "2020-07-03", "2020-09-07",     "2020-11-26", "2020-12-25", "2021-01-01", "2021-01-18", "2021-02-15", "2021-04-02", "2021-05-31", "2021-06-18", "2021-07-05", "2021-09-06", "2021-11-25", "2021-12-24", "2022-01-17", "2022-02-21", "2022-04-15", "2022-05-30", "2022-06-20", "2022-07-04"))) 
          #Removing holiday dates from final dates vector
          finaldates <- anti_join(dates, holidaydates)
            #Combining data and market open dates
            finaldata <- as_tibble(cbind(parsed_tibble, finaldates))
  
  #Change column names
  colnames(finaldata) <- c("volume", "volume_weighted", "open", "close", "high", "low", "Unix_Msec", "transactions", "date")
  
  #Return tibble of data
  return(finaldata)
}
```

# **Data Exploration**

This portion fo the vignette will now use the functions for interacting with the Polygon Financial API to retrieve some data and look at some of the endpoints. 

The S&P Oil & Gas Exploration & Production Industry Index is comprised of 61 constituents. In this, we'll just take a look at one of the biggest common stock contributors, Valero Energy Corporation (VLO), to see generally how the oil industry has been doing. Additionally, in the latter part of this section, we'll take a look at 2 ETF's that follow this industry index. DRIP is a 2x leveraged *inverse* ETF, meaning DRIP returns 2x the inverse of the daily oil industry index performance. Conversely, GUSH tracks with the index and is a standard 2x leveraged ETF. Using data from the api, we'll look into some basic stock endpoints, look at trends since 2020, and show some differences between VLO and those ETF's. 

Let's start with some simple application of the market cap function. This function queries the polygon api and returns the ticker of your search, marketcap, and year of you search. Here, I'm going to get the market cap value for the VLO ticker in 2020, 2021 and 2022 (in June of each year). Then, I'm going to make a bar plot of the market caps by year to look for any trends. 

VLO:
```{r, echo = TRUE, eval = TRUE}
#Getting market cap for 2020 (non-weekend)
VLOcap_20 <- marketcap("VLO", "2020-06-26")
  #Getting marketcap for 2021
  VLOcap_21 <- marketcap("VLO", "2021-06-25")
    #Getting market cap for 2022
    VLOcap_22 <- marketcap("VLO", "2022-06-24")

#Putting these caps together to create a dataset to plot from
VLOcaps <- as_tibble(bind_rows(VLOcap_20, VLOcap_21, VLOcap_22))

#Changing market cap to a numeric value for better plotting
VLOcaps$marketcap <- as.numeric(VLOcaps$marketcap)

#Plotting market cap for VLO over past 2 years
plota <- ggplot(VLOcaps, aes(x = ticker, y = marketcap)) +
  #Create side-by-side bar plot filled by year
  geom_bar(aes(fill=year), stat = "identity", position = "dodge") +
  #Add title and labels
  labs( x = "Ticker", y = "Market Cap", title = "VLO Market Cap by Year") +
  #Set legend title
  scale_fill_discrete(name = "Year") 

plota
```

Okay, so the marketcap function works in allowing us to retrieve market cap data on a common stock oil company. Plotting those values in a bar plot by year let us see that Valero Oil has increased market cap significantly each year since 2020. 

Now, let's switch gears and use the aggregate function to retrieve some data from the oil industry regarding stock price and stock volume. First, we'll take a look at Valero again to explore price and volume, then take a look at the 2 popular leveraged ETF's of the index: DRIP (2x inverse) and GUSH (2x), to see if the ETF's follow a different trend than the common stocks. Since DRIP is an inverse, we should expect the price trends to be inverse that of all other stocks. 

Let's start by exploring VLO:
```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
VLO <- aggregate("VLO")

#Creating avg daily price column
VLO$avgprice <- ((VLO$high + VLO$low)/2)

#Quick plot of average daily price over the past 2 years
plotVLO <- ggplot(VLO, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth() +
  labs( x = "Date", y = "Avg Daily Price", title = "VLO Average Daily Price Data")

plotVLOv <- ggplot(VLO, aes(x=date, y=volume)) +
  geom_point() +
  labs( x = "Date", y = "Volume", title = "VLO Volume Data")

plotVLO
plotVLOv
```

Valero has shown a solid increase in average daily stock price and consistent volume across the price increase. Safe to say they have been doing well the past two years!

Now, let's look at those leveraged ETF's, starting with the inverse, DRIP:
```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
DRIP <- aggregate("DRIP")

#Creating avg daily price column
DRIP$avgprice <- ((DRIP$high + DRIP$low)/2)

#Plot of average daily price over the past 2 years
plotDRIP <- ggplot(DRIP, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth() +
  labs( x = "Date", y = "Avg Daily Price", title = "DRIP Average Daily Price Data")

#Plot of daily volume over the past 2 years
plotDRIPv <- ggplot(DRIP, aes(x=date, y=volume)) +
  geom_point() +
  labs( x = "Date", y = "Volume", title = "DRIP Volume Data")

plotDRIP
plotDRIPv
```

Well, that looks different! As expected, the inverse ETF has a trend opposite the oil industry index stocks. However, what stands out to me is the volume scatter plot. Volume seems to have gone hyperbolic in the year 2022, perhaps because price is so low and buyers expect a pull back from the impressive increase in the oil industry. 

Let's see about GUSH:
```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
GUSH <- aggregate("GUSH")

#Creating avg daily price column
GUSH$avgprice <- ((GUSH$high + GUSH$low)/2)

#Quick plot of average daily price over the past 2 years
plotGUSH <- ggplot(GUSH, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth() +
  labs( x = "Date", y = "Avg Daily Price", title = "GUSH Average Daily Price Data")


plotGUSHv <- ggplot(GUSH, aes(x=date, y=volume)) +
  geom_point() +
  labs( x = "Date", y = "Volume", title = "GUSH Volume Data")

plotGUSH
plotGUSHv
```

As expected, since GUSH follows the industry, the price trend looks very similar to that of the Big 3. However, volume has decreased over time significantly more than Valero. It certainly seems like the tide is turning away from a bullish oil market and more towards a bearish one.















