---
title: "Using APIs: Creating Vignette on Stock Data as Example"
author: "Evan Brown"
date: '2022-06-23'
output: 
  github_document:
    toc: true
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "../images/")
```

This is a vignette to show how to retrieve data from an [API](https://aws.amazon.com/what-is/api/#:~:text=API%20stands%20for%20Application%20Programming,other%20using%20requests%20and%20responses.). To demonstrate, I'll be working with the [Polygon Financial API](https://polygon.io/docs/stocks/getting-started). For the purposes of this work, I used a free account with Polygon, limiting my data retrieval "from-date" to 2 years. 

In this vignette, I will work to pull data from the [S&P Oil & Gas Exploration & Production Industry Index](https://www.spglobal.com/spdji/en/indices/equity/sp-oil-gas-exploration-production-select-industry-index/#data) and compare endpoints of some top index constituents to industry [ETF's](https://www.investopedia.com/terms/e/etf.asp#:~:text=An%20exchange%2Dtraded%20fund%20(ETF)%20is%20a%20type%20of,that%20a%20regular%20stock%20can.) and see if I can find any similarities or differences over the past 2 years. 


# **Required Functions**

These R packages are required in order to work with the Polygon API:

- [tidyverse](https://www.tidyverse.org/packages/)  
- [jsonlite](https://cran.r-project.org/web/packages/jsonlite/index.html) 
- [httr](https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html)  

```{r, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
#Reading in packages
library(tidyverse)
library(httr)
library(jsonlite)
```

# **Functions for API Interaction**

This portion of the vignette will show where I define functions for interacting with the Polygon Financial API.

## 'aggregate' function

This function returns daily stock endpoints from 06-21-2021 to 06-22-2022 as a tibble for the user provided tibble. apikey default is my key, but users of the function should pass their own key in the function call options. User keys can be generated [here](https://polygon.io/).

```{r, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

aggregate<- function(ticker, apikey = "OrlbxnjeCyqGDGkKtpIqxKKs0f8Eh77C"){
  
  #Set URL partitions
  baseURL <- "https://api.polygon.io/v2/aggs/ticker/"
    ticker <- ticker
      adjustments <- "/range/1/day/2020-06-24/2022-06-24?adjusted=true&sort=asc&limit=5000&apiKey="
        apikey <- apikey
  
  #Paste partitions to get full URL with desired ticker
  fullURL <- paste0(baseURL, ticker, adjustments, apikey)
  
  #Get api data using get function
  api_data <- GET(fullURL)
    #Convert to character string with fromJSON function
    parsed_api_data <- fromJSON(rawToChar(api_data$content))
      #Save as R object
      parsed_api_data_results <- parsed_api_data$results
        #Convert to tibble
        parsed_tibble <- as_tibble(parsed_api_data_results)

  #Create sequence of raw dates
  dates <- (seq(as.Date("2020-06-24"),as.Date("2022-06-24"),by = 1))
    #Filtering weekends (market not open)
    dates <- as.tibble((dates[!weekdays(dates) %in% c("Saturday","Sunday")]))
      #Filtering holidays (market not open)
        holidaydates <- as.tibble(as.Date(c("2020-01-01", "2020-01-20", "2020-02-17", "2020-04-10", "2020-05-25", "2020-07-03", "2020-09-07",     "2020-11-26", "2020-12-25", "2021-01-01", "2021-01-18", "2021-02-15", "2021-04-02", "2021-05-31", "2021-06-18", "2021-07-05", "2021-09-06", "2021-11-25", "2021-12-24", "2022-01-17", "2022-02-21", "2022-04-15", "2022-05-30", "2022-06-20", "2022-07-04"))) 
          #Removing holiday dates from final dates vector
          finaldates <- anti_join(dates, holidaydates)
            #Combining data and market open dates
            finaldata <- as_tibble(cbind(parsed_tibble, finaldates))
  
  #Change column names
  colnames(finaldata) <- c("volume", "volume_weighted", "open", "close", "high", "low", "Unix_Msec", "transactions", "date")
  
  #Return tibble of data
  return(finaldata)
}
```

## 'marketcap' Function

This function returns the [market cap](https://www.investopedia.com/terms/m/marketcapitalization.asp) for the user provided ticker and date. apikey default is my key, but users of the function should pass their own key in the function call options. User keys can be generated [here](https://polygon.io/).

```{r, echo=TRUE, eval=FALSE, warning=FALSE, message=FALSE}

marketcap <- function(ticker, date, apikey = "OrlbxnjeCyqGDGkKtpIqxKKs0f8Eh77C"){
  
  #Set URL partitions
  baseURL <- "https://api.polygon.io/v3/reference/tickers/"
    ticker <- ticker
      dateadjustment <- "?date="
        date <- date
          adjustments <- "&apiKey="
            apikey <- apikey
          
  #Paste partitions to get full URL with desired ticker
  fullURL <- paste0(baseURL, ticker, dateadjustment, date, adjustments, apikey)   
  
  #Get api data using get function
  api_data <- GET(fullURL)
    #Convert to character string with fromJSON function
    parsed_api_data <- fromJSON(rawToChar(api_data$content))
      #Save as R object
      parsed_api_data_results <- parsed_api_data$results
      results <- as_data_frame(parsed_api_data_results[names(parsed_api_data_results) != "address"])
      selecteddata <- results %>% select(1, 12) 
      finaldata <- distinct(selecteddata)
      
      return(finaldata)

}

```


# **Data Exploration**

This portion fo the vignette will now use the function for interacting with the Polygon Financial API to retrieve some data and look at some of the endpoints. 

The S&P Oil & Gas Exploration & Production Industry Index is comprised of 61 constituents. Let's look 3 of the biggest contributors: Valero Energy Corp (VLO), Marathon Oil Corp (MRO), and Hess Corp (HES). Here, I'll work with the marketcap function to retrieve the market cap data from June 1, 2022 and June 1, 2021, then organize it into a data set that we can plot from. 
```{r, echo = TRUE, eval = FALSE, warning=FALSE, message=FALSE}

VLOcap_21 <- marketcap("VLO", "2021-06-01") 
  VLOcap_22 <- marketcap("VLO", "2022-06-01")
    MROcap_21 <- marketcap("MRO", "2021-06-01")
      MROcap_22 <- marketcap("MRO", "2022-06-01")
        HEScap_21 <- marketcap("HES", "2021-06-01")
          HEScap_22 <- marketcap("HES", "2022-06-01")

#Create master dataset with marketcaps for all 3 companies in 2021 and 2022
Bigoil_marketcaps <- as_tibble(bind_rows(VLOcap_21, VLOcap_22, MROcap_21, MROcap_22, HEScap_21, HEScap_22))

#Lets change marketcap to a numeric value for better plotting
Bigoil_marketcaps$marketcap <- as.numeric(Bigoil_marketcaps$marketcap)
```

Alright, now let's plot that data and see what it looks like. 
```{r, echo = TRUE, eval = FALSE, warning=FALSE, message=FALSE}

plot1 <- ggplot(Bigoil_marketcaps, aes(x = ticker, y = marketcap)) +
  geom_bar(aes(fill=date), stat = "identity", position = "dodge") +
  labs( x = "Ticker", y = "Market Cap", title = "Plot of Market Cap for Oil Index Big 3 for 2021 & 2022") +
  scale_fill_discrete(name = "Year", labels = c("2021", "2022")) 

plot1
```

Okay, so clearly an increase in market cap for the oil industry "Big 3" from 2021 to 2022. Since market cap is indicative of the overall value of the companies, it's safe to say these guys have done well the past year. 


**Let's use the aggregate function to retrieve some data from the oil industry to explore recent price trends. First, we'll take a look at the oil industry "Big 3" then look at 2 leveraged ETF's of the index: DRIP (2x bear) and GUSH (2x bull). **

Let's start with exploring VLO.

```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
VLO <- aggregate("VLO")

#Creating avg daily price column
VLO$avgprice <- ((VLO$high + VLO$low)/2)

#Quick plot of average daily price over the past 2 years
plot3 <- ggplot(VLO, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Avg Daily Price", title = "VLO Average Daily Price Data")
plot3

plot4 <- ggplot(VLO, aes(x=date, y=volume)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Volume", title = "VLO Volume Data")
plot4
```

Now, MRO.

```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
MRO <- aggregate("MRO")

#Creating avg daily price column
MRO$avgprice <- ((MRO$high + MRO$low)/2)

#Quick plot of average daily price over the past 2 years
plot5 <- ggplot(MRO, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Avg Daily Price", title = "MRO Average Daily Price Data")
plot5

plot6 <- ggplot(MRO, aes(x=date, y=volume)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Volume", title = "MRO Volume Data")
plot6
```

Now, HES.

```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
HES <- aggregate("HES")

#Creating avg daily price column
HES$avgprice <- ((HES$high + HES$low)/2)

#Quick plot of average daily price over the past 2 years
plot7 <- ggplot(HES, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Avg Daily Price", title = "HES Average Daily Price Data")
plot7

plot8 <- ggplot(HES, aes(x=date, y=volume)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Volume", title = "HES Volume Data")
plot8
```

All of the oil industry Big 3 companies show a solid increase in average daily stock price and consistent volume across the price increase. Safe to say these companies have been doing well the past two years!


Now, let's explore two popular leveraged ETF's for the oil and gas industry. DRIP is a 2x leveraged inverse ETF, and GUSH is a standard 2x leveraged ETF. 

```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
DRIP <- aggregate("DRIP")

#Creating avg daily price column
DRIP$avgprice <- ((DRIP$high + DRIP$low)/2)

#Quick plot of average daily price over the past 2 years
plot9 <- ggplot(DRIP, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Avg Daily Price", title = "DRIP Average Daily Price Data")
plot9

plot10 <- ggplot(DRIP, aes(x=date, y=volume)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Volume", title = "DRIP Volume Data")
plot10
```

Well, that looks different! DRIP clearly hasn't followed the same moderate trends. Let's see what GUSH looks like.

```{r, echo = TRUE, eval = TRUE, warning=FALSE, message=FALSE}
#Retrieving data on DRIP ticker
GUSH <- aggregate("GUSH")

#Creating avg daily price column
GUSH$avgprice <- ((GUSH$high + GUSH$low)/2)

#Quick plot of average daily price over the past 2 years
plot11 <- ggplot(GUSH, aes(x=date, y=avgprice)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Avg Daily Price", title = "GUSH Average Daily Price Data")
plot11

plot12 <- ggplot(GUSH, aes(x=date, y=volume)) +
  geom_point() +
  geom_smooth(method = gamma) +
  labs( x = "Date", y = "Volume", title = "GUSH Volume Data")
plot12
```

As expected, since GUSH follows the industry, the price trend looks very similar to that of the Big 3. However, volume decreases over time significantly more than those Big 3 companies. 















